---
title: "2025_11_24_Supplementary_analysis_associate_mutational_burden_with_age"
output: html_document
date: "2025-11-24"
---

**Purpose:**

* Test for an association of mutational burden with age within immunocompetent and immunosuppressed patients

**Methods:**

* Mutational burden quantified

* Associate mutational burden with age

* Subset approximate mutational burdens by signature.

**Results:**

## Libraries
```{r}
library(ggplot2)
library(maftools)
library(BSgenome.Hsapiens.UCSC.hg38)
library(deconstructSigs)
library(sigminer)
library(stringr)
library(ggsci)
library(plyr)
library(reshape2)
library(dunn.test)
```


## Read in Mayo variants

```{r}
# Read in mayo file
maf_file_Mayo = "~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/MAF_file/Mayo_data_with_new_samples_FFPE_fitlered_DNP_fixed.maf"
maf_mayo <- data.table::fread(maf_file_Mayo)
maf_mayo <- maf_mayo[which((maf_mayo$t_alt_count)>=5),]
maf_mayo <- maf_mayo[which((maf_mayo$t_depth)>=10),]
readr::write_tsv(maf_mayo, "~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/MAF_file/Mayo_data_with_new_samples_FFPE_fitlered_coverage_filtered.maf")
maf_file_Mayo = "~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/MAF_file/Mayo_data_with_new_samples_FFPE_fitlered_coverage_filtered.maf"
table(maf_mayo$Variant_Classification)
vcNames <- names(maf_mayo$Variant_Classification)
variants_mayo = read.maf(maf = maf_file_Mayo, vc_nonSyn=vcNames)

# Calculate mutational burden
mutational_burden_mayo <- variants_mayo@variant.classification.summary
maf_mayo$number <- 1
total_mutations_mayo <- ddply(maf_mayo, "Tumor_Sample_Barcode", numcolwise(sum))
total_mutations_mayo <- total_mutations_mayo[match(mutational_burden_mayo$Tumor_Sample_Barcode, total_mutations_mayo$Tumor_Sample_Barcode),]
identical(as.character(mutational_burden_mayo$Tumor_Sample_Barcode), total_mutations_mayo$Tumor_Sample_Barcode)
mutational_burden_mayo$somatic_mutations <- total_mutations_mayo$number

### Add immunosuppression annotations
clinical_data <- as.data.frame(read.csv("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Metadata/Metadata_annotations_all_Mayo_samples.csv"))
clinical_data <- clinical_data[which(clinical_data$Include_in_study=="Yes"),]
clinical_data <- clinical_data[-which(is.na(clinical_data$Tumor_Sample_Barcode)),]
#clinical_data <- clinical_data[-which(is.na(clinical_data$RNA_ID)),]
proportion_binding <- as.data.frame(read.csv("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Metadata/Metadata_annotations_all_Mayo_samples_immune_binary.csv"))
proportion_binding_matched <- proportion_binding[match(clinical_data$Tumor_Sample_Barcode, proportion_binding$Tumor_Sample_Barcode),]
clinical_data$immune_binary <- proportion_binding_matched$immune_binary
clinical_data[which(is.na(clinical_data$immune_binary) & clinical_data$Immunosuppression=="No"),18] <- "No RNA"
clinical_data[which(is.na(clinical_data$immune_binary) & clinical_data$Immunosuppression=="Yes"),18] <- "Suppressed"
clinical_data <- clinical_data[which(clinical_data$Dataset=="Old"),]
table(clinical_data$immune_binary)

mutational_burden_mayo <- mutational_burden_mayo[which(mutational_burden_mayo$Tumor_Sample_Barcode %in% clinical_data$Tumor_Sample_Barcode),]
clinical_data <- clinical_data[match(mutational_burden_mayo$Tumor_Sample_Barcode, clinical_data$Tumor_Sample_Barcode),]
identical(as.character(mutational_burden_mayo$Tumor_Sample_Barcode), clinical_data$Tumor_Sample_Barcode)
mutational_burden_mayo$immunosuppression <- clinical_data$immune_binary
mutational_burden_mayo$met <- clinical_data$Met
mutational_burden_mayo$stage <- clinical_data$Stage
mutational_burden_mayo$sex <- clinical_data$Gender
mutational_burden_mayo$age <- clinical_data$Age
mutational_burden_mayo$treatment <- clinical_data$Immunosuppressive_agent
mutational_burden_mayo$location <- clinical_data$Location
mutational_burden_mayo$Dataset <- clinical_data$Dataset

mutational_burden <- mutational_burden_mayo
total_data <- mutational_burden[,c(1,3,9,10,11,12,13,14,15,16,17)]
```

## Mutational burden plots

```{r}
mutational_burden_formatted <- total_data[,c(1:7,9:10)]
mutational_burden_formatted <- reshape2::melt(mutational_burden_formatted, value.name = "Mutations")
somatic_burden_matched <- total_data[match(mutational_burden_formatted$Tumor_Sample_Barcode, total_data$Tumor_Sample_Barcode),]
identical(mutational_burden_formatted$Tumor_Sample_Barcode, somatic_burden_matched$Tumor_Sample_Barcode)
mutational_burden_formatted$somatic_mutations <- somatic_burden_matched$somatic_mutations
mutational_burden_formatted$age <- somatic_burden_matched$age

ggplot(mutational_burden_formatted, aes(x = age, y= somatic_mutations, color= immunosuppression)) + geom_point() + geom_smooth(method= "lm", se=FALSE) + theme_minimal()
ggplot(mutational_burden_formatted, aes(x = age, y= somatic_mutations)) + geom_point() + geom_smooth(method= "lm") + theme_minimal()

high <- mutational_burden_formatted[which(mutational_burden_formatted$immunosuppression == "high"),]
model <- lm(somatic_mutations ~ age, high)
summary(model)

low <- mutational_burden_formatted[which(mutational_burden_formatted$immunosuppression == "low"),]
model <- lm(somatic_mutations ~ age, low)
summary(model)

suppressed <- mutational_burden_formatted[which(mutational_burden_formatted$immunosuppression == "Suppressed"),]
model <- lm(somatic_mutations ~ age, suppressed)
summary(model)

## Test plot without outliers

mutational_burden_formatted_lim <- mutational_burden_formatted[-which(mutational_burden_formatted$Tumor_Sample_Barcode=="s_CSCC_18_DNA"),]
mutational_burden_formatted_lim <- mutational_burden_formatted_lim[-which(mutational_burden_formatted_lim$Tumor_Sample_Barcode=="s_CSCC_48_DNA"),]
ggplot(mutational_burden_formatted_lim, aes(x = age, y= somatic_mutations, color= immunosuppression)) + geom_point() + geom_smooth(method= "lm", se=FALSE) + theme_minimal()
high <- mutational_burden_formatted[which(mutational_burden_formatted_lim$immunosuppression == "high"),]
model <- lm(somatic_mutations ~ age, high)
summary(model)
```

## Subset by mutation type with age

```{r}
filtered_sig <- as.data.frame(read.csv("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Mutational burden Figures/sigminer_signatures.csv"))
row.names(filtered_sig) <- filtered_sig[,1]
filtered_sig <- filtered_sig[,-1]

filtered_sig <- t(filtered_sig)
row.names(filtered_sig) <- gsub("\\.", "-", row.names(filtered_sig))
filtered_sig <- filtered_sig[match(mutational_burden_formatted$Tumor_Sample_Barcode, row.names(filtered_sig)),]
identical(as.character(mutational_burden_formatted$Tumor_Sample_Barcode), row.names(filtered_sig))

mutational_burden_formatted <- cbind(mutational_burden_formatted, filtered_sig)
mutational_burden_formatted$UV_sig <- as.numeric(mutational_burden_formatted$Sig1) + as.numeric(mutational_burden_formatted$Sig10) + as.numeric(mutational_burden_formatted$Sig2)
```

## Plot UV and clock-like signature by age

```{r}
ggplot(mutational_burden_formatted, aes(x = age, y= UV_sig* somatic_mutations, color= immunosuppression)) + geom_point() + geom_smooth(method= "lm") + theme_minimal()

suppressed <- mutational_burden_formatted[which(mutational_burden_formatted$immunosuppression == "Suppressed"),]
model <- lm((somatic_mutations*UV_sig) ~ age, suppressed)
summary(model)

ggplot(mutational_burden_formatted, aes(x = age, y= (as.numeric(Sig3))* somatic_mutations, color= immunosuppression)) + geom_point() + geom_smooth(method= "lm") + theme_minimal()

outlier_removed <- mutational_burden_formatted[-which(mutational_burden_formatted$Sig3 > 0.5),]
ggplot(outlier_removed, aes(x = age, y= (as.numeric(Sig3))* somatic_mutations, color= immunosuppression)) + geom_point() + geom_smooth(method= "lm") + theme_minimal()
ggplot(outlier_removed, aes(x = age, y= (as.numeric(Sig3))* somatic_mutations)) + geom_point() + geom_smooth(method= "lm") + theme_minimal()
```