---
title: "2025_2_18_Compare_enrichment_In_silico_observed"
output: html_document
date: '2025-02-18'
---

# **Purpose:**

Compare the enrichment of neoantigens at low mRNA to scrambled HLA types, using new methods from Bonnie and 

# **Methods:**

* Calculate strong binding and weak binding neoantigens for both patient-specific and scrambled HLA types

* Assign each an expression value

* Rank order expression values

* Sum ranks of predicted binding neoantigens and all other neoantigens

* Possible: divide sums by the total number of predicted binding neoantigens and total mutations

* Calculate the difference in rank sums between the predicted binding neoantigens and total mutations

# **Results:**

### Libraries
```{r, message=FALSE, warning=FALSE}
library(DGEobj.utils)
library(ggplot2)
library(beanplot)
library(stringr)
```


### Read in in silico data

```{r}
clinical_data <- as.data.frame(read.csv("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Characteristics clonal subclonal neoags/Compiled_in_silico_enrichment.csv"))

maf_file = "~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/MAF_file/Mayo_data_with_new_samples_FFPE_fitlered_coverage_filtered.maf"
MAF <- data.table::fread(maf_file)

expression <- as.data.frame(read.csv("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/RNA_seq/star/all_samples_tpm.counts.txt", row.names = 1))
colnames(expression) <- gsub("NPT\\.", "NPT-", colnames(expression))
colnames(expression) <- gsub("\\.1", "-1", colnames(expression))
colnames(expression) <- gsub("\\.2", "-2", colnames(expression))
genes <- rownames(expression)
```


```{r}
for (a in 1:length(clinical_data$Patient)){
  print(a)
  ### Define IDs
  sample <- clinical_data$Tumor_Sample_Barcode[a]
  patient <- clinical_data$ID[a]
  rna <- clinical_data$RNA_ID[a]

  ### Subset MAF file
  MAF_subset <- as.data.frame(MAF[which(MAF$Tumor_Sample_Barcode==sample),])
  MAF_subset <- MAF_subset[, which(colnames(MAF_subset) %in% c("Chromosome", "Start_Position", "Hugo_Symbol", "Reference_Allele",
                                                            "Tumor_Seq_Allele2","t_ref_count", "t_alt_count"))]

  ### Create Mutation ID
  MAF_subset$ID <- paste(MAF_subset$Chromosome, MAF_subset$Start_Position-1, MAF_subset$Reference_Allele, MAF_subset$Tumor_Seq_Allele2, sep=":")
  MAF_subset$ID_lim <- paste(MAF_subset$Chromosome, MAF_subset$Start_Position-1, sep=":")

  ### Match Expression
  expression_subset <- as.data.frame(expression[,which(colnames(expression)== rna)])
  expression_subset$genes <- genes
  colnames(expression_subset) <- c("TPM","Gene")
  
  expression_subset <- expression_subset[match(MAF_subset$Hugo_Symbol, expression_subset$Gene),]
  identical(MAF_subset$Hugo_Symbol[10], expression_subset$Gene[10])
  
  MAF_subset$Expression <- expression_subset$TPM
  
  ### Match Peptides
  peptides <- as.data.frame(read.table(paste("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/MT_peptides/", patient, ".somaticvariants_filtered_vep.peptides", sep=""), header=FALSE))
  MAF_subset <- MAF_subset[which(MAF_subset$ID %in% peptides$V5),]
  peptides <- peptides[match(MAF_subset$ID, peptides$V5),]
  identical(MAF_subset$ID, peptides$V5)
  MAF_subset$Peptide <- peptides$V6

  ### Limit MAF columns
  MAF_subset <- MAF_subset[,-c(1:6)]
  
  ### Read in binding and remove WT
  binding <- as.data.frame(read.table(paste("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/netMHCpan/", patient, "_netmhc_9_10mer.xsl", sep=""), skip=1, header=TRUE))
  
  wt_all <- as.data.frame(read.csv(paste("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/WT_peptides/", patient, ".WT_n_mers_9_10_15mer.peptides", sep=""), header=FALSE))
  binding <- binding[-which(binding$Peptide %in% wt_all$V2),]
  binding <- unique(binding)
  IDs <- stringr::str_split_fixed(binding$ID, "_", 4)
  IDs[,1] <- paste("chr", IDs[,1], sep="")
  binding$ID <- paste(IDs[,1], IDs[,2], sep=":")
  

  MAF_subset$weak_neo <- "not+neo"
  
  ### Match min binding
  for (i in 1:length(MAF_subset$ID)){
    #print(i)
    binding_subset <- binding[which(binding$ID==MAF_subset$ID_lim[i]),]
    MAF_subset$binding[i] <- min(binding_subset$nM, binding_subset$nM.1, binding_subset$nM.2, binding_subset$nM.3, binding_subset$nM.4, binding_subset$nM.5)
    if (MAF_subset$binding[i] <500){
      MAF_subset$weak_neo[i] <- "neo"
    }
  }

  if (length(MAF_subset[-which(is.na(MAF_subset$Expression)),1]>0)){
   MAF_subset <- MAF_subset[-which(is.na(MAF_subset$Expression)),] 
  }
  MAF_subset$Sample <- sample
  
  MAF_subset_rank_order <- MAF_subset[order(as.numeric(MAF_subset$Expression)),]
  MAF_subset_rank_order$rank <- dplyr::dense_rank(as.numeric(MAF_subset_rank_order$Expression))
  
  ## True binders
  not_neo <- MAF_subset_rank_order[which(MAF_subset_rank_order$weak_neo=="not+neo"),]
  not_neo_sum <- sum(not_neo$rank)/length(not_neo$ID)
  neo <- MAF_subset_rank_order[which(MAF_subset_rank_order$weak_neo=="neo"),]
  neo_sum <- sum(neo$rank)/length(neo$ID)
  clinical_data$weak_enrichment[a] <- not_neo_sum - neo_sum
}
```

## Plot all data no split

```{r}
clinical_data_plot <- clinical_data[,which(colnames(clinical_data) %in% c("Patient", "Immunosuppression", "weak_enrichment", "in_silico_enrichment"))]
clinical_data_plot <- reshape2::melt(clinical_data_plot)

ggplot(clinical_data_plot, aes(x=variable, y=value)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()

wilcox.test(clinical_data$weak_enrichment, clinical_data$in_silico_enrichment, paired=TRUE)
```

## Plot all data split

```{r}
immune_binary <- as.data.frame(read.csv("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Metadata/Metadata_annotations_all_Mayo_samples_immune_binary.csv"))
immune_binary <- immune_binary[match(clinical_data$ID, immune_binary$ID),]
identical(clinical_data$ID, immune_binary$ID)
clinical_data$immune_binary <- immune_binary$immune_binary

clinical_data_plot <- clinical_data[,which(colnames(clinical_data) %in% c("Patient", "immune_binary", "weak_enrichment", "in_silico_enrichment"))]
clinical_data_plot <- reshape2::melt(clinical_data_plot)
clinical_data_plot$new_variable <- paste(clinical_data_plot$immune_binary, clinical_data_plot$variable)

ggplot(clinical_data_plot, aes(x=new_variable, y=value)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()

pdf("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Characteristics clonal subclonal neoags/Comparison_in_silico_figures/Enrichment_comparison.pdf", width=8, height=5)
ggplot(clinical_data_plot, aes(x=new_variable, y=value)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + theme_minimal()
dev.off()

high <- clinical_data[which(clinical_data$immune_binary=="high"),]
wilcox.test(high$weak_enrichment, high$in_silico_enrichment, paired=TRUE)
low <- clinical_data[which(clinical_data$immune_binary=="low"),]
wilcox.test(low$weak_enrichment, low$in_silico_enrichment, paired=TRUE)
suppressed <- clinical_data[which(clinical_data$immune_binary=="Suppressed"),]
wilcox.test(suppressed$weak_enrichment, suppressed$in_silico_enrichment, paired=TRUE)

test <- clinical_data[which(clinical_data$immune_binary=="low" | clinical_data$immune_binary=="high"),]
wilcox.test(test$weak_enrichment, test$in_silico_enrichment, paired=TRUE)
```

