---
title: "2025_12_10_Map_Clusters_HiTide_data"
output: html_document
date: "2025-12-10"
---

## **Purpose:**

* Test the mapping of the immunogenic/non-immunogenic peptides from HiTide only the clusters

## **Methods:**

* Read in the existing kmeans clusters

* Format new data to match with appropriate normalization

* Visualize mapping

* Calculate proportion of immunogenic and non-immunogenic neoantigens that map to each cluster

### Libraries
```{r}
library(flexclust)
library(pheatmap)
library(stringr)
library(ggplot2)
library(ggpubr)
library(biomaRt)
library(DGEobj.utils)
library(dplyr)
library(readr)
library(tidyr)
library(tibble)
```

### Read in existing model and convert to kcca model
```{r}
kmeans_model <- readRDS("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Characteristics clonal subclonal neoags/Clustering_analysis/map_hitide/kmeans_model.rds")

original_data <- as.data.frame(read.csv("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Characteristics clonal subclonal neoags/Clustering_analysis/map_hitide/data_clust.csv"))
head(original_data)

cl1 <- as.kcca(kmeans_model, original_data)
cl1
```

### Format new data to match original data

```{r}
new_data <- as.data.frame(read.csv("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Subcellular_localization/HiTide_data/HiTide_subset_data_characteristics_add_EC.txt"))
new_data$DAI <- new_data$netMHC_nM_WT/new_data$netMHC_nM_MT
new_data_rearranged <- new_data[,c(9,10,12,8,7,14,15)]
colnames(new_data_rearranged)
colnames(new_data_rearranged) <- colnames(original_data)
new_data_rearranged$dataset <- new_data$dataset
new_data_rearranged$immunogenic <- new_data$response_type
```

### Scale new data to match original data
```{r}
new_data_rearranged$expression <- scale(log(new_data_rearranged$expression+0.01,10), center = TRUE)
new_data_rearranged$TAP <- scale(new_data_rearranged$TAP, center=TRUE)
new_data_rearranged$Cle <- scale(new_data_rearranged$Cle, center=TRUE)
new_data_rearranged$min_binding <- scale(log(new_data_rearranged$min_binding,10), center=TRUE)
new_data_rearranged$max_stability <- scale(log(as.numeric(new_data_rearranged$max_stability),10), center=TRUE)
new_data_rearranged$DAI <- scale(log(new_data_rearranged$DAI+0.01,10), center=TRUE)
new_data_rearranged$EC <- scale(new_data_rearranged$EC, center=TRUE)
```

### Predict for new data
```{r}
set.seed(123)
new_data_rearranged$pred <- predict(cl1, new_data_rearranged[,1:7])
```

### Plot proportion in each cluster
```{r}
table(new_data_rearranged$immunogenic)
table(new_data_rearranged$pred, new_data_rearranged$immunogenic)
table <- as.data.frame(table(new_data_rearranged$pred, new_data_rearranged$immunogenic))
table[1:7,3] <- table[1:7,3]/151
table[8:14,3] <- table[8:14,3]/3788
ggplot(table, aes(x=Var1, y=Freq, fill=Var2)) + geom_bar(stat="identity", position="dodge") + theme_minimal()

pdf("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Subcellular_localization/HiTide_data/Proportion_each_cluster.pdf")
ggplot(table, aes(x=Var1, y=Freq, fill=Var2)) + geom_bar(stat="identity", position="dodge") + theme_minimal()
dev.off()
```

### Plot proportion in each cluster subset by dataset
```{r}
hitide_subset <- new_data_rearranged[which(new_data_rearranged$dataset == "HiTIDE"),]
table(hitide_subset$immunogenic)
table(hitide_subset$pred, hitide_subset$immunogenic)
table <- as.data.frame(table(hitide_subset$pred, hitide_subset$immunogenic))
table[1:7,3] <- table[1:7,3]/30
table[8:14,3] <- table[8:14,3]/1201
ggplot(table, aes(x=Var1, y=Freq, fill=Var2)) + geom_bar(stat="identity", position="dodge") + theme_minimal()

pdf("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Subcellular_localization/HiTide_data/HiTide_Proportion_each_cluster.pdf")
ggplot(table, aes(x=Var1, y=Freq, fill=Var2)) + geom_bar(stat="identity", position="dodge") + theme_minimal()
dev.off()

nci_subset <- new_data_rearranged[which(new_data_rearranged$dataset == "NCI"),]
table(nci_subset$immunogenic)
table(nci_subset$pred, nci_subset$immunogenic)
table <- as.data.frame(table(nci_subset$pred, nci_subset$immunogenic))
table[1:7,3] <- table[1:7,3]/90
table[8:14,3] <- table[8:14,3]/2000
ggplot(table, aes(x=Var1, y=Freq, fill=Var2)) + geom_bar(stat="identity", position="dodge") + theme_minimal()

pdf("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Subcellular_localization/HiTide_data/NCI_Proportion_each_cluster.pdf")
ggplot(table, aes(x=Var1, y=Freq, fill=Var2)) + geom_bar(stat="identity", position="dodge") + theme_minimal()
dev.off()

tesla_subset <- new_data_rearranged[which(new_data_rearranged$dataset == "TESLA"),]
table(tesla_subset$immunogenic)
table(tesla_subset$pred, tesla_subset$immunogenic)
table <- as.data.frame(table(tesla_subset$pred, tesla_subset$immunogenic))
table[1:7,3] <- table[1:7,3]/31
table[8:14,3] <- table[8:14,3]/587
ggplot(table, aes(x=Var1, y=Freq, fill=Var2)) + geom_bar(stat="identity", position="dodge") + theme_minimal()

pdf("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Subcellular_localization/HiTide_data/TESLA_Proportion_each_cluster.pdf")
ggplot(table, aes(x=Var1, y=Freq, fill=Var2)) + geom_bar(stat="identity", position="dodge") + theme_minimal()
dev.off()
```

### Fisher exact test
```{r}
table <- as.data.frame(table(new_data_rearranged$pred, new_data_rearranged$immunogenic))
df <- table %>%
      # Convert long data into contigency table
      pivot_wider(names_from = Var1,   
                  values_from = Freq,    
                  values_fill = 0 ) %>%
      column_to_rownames("Var2") %>%
      as.matrix()

df
fisher.test(df, simulate.p.value = TRUE)
df <- t(df)
fisher.test(df, simulate.p.value = TRUE)
```

### Residuals
```{r}
chi <- chisq.test(df)
res <- as.data.frame(as.table(chi$residuals))
names(res) <- c("Cluster", "Immunogenic", "StdResid")

ggplot(res, aes(x = factor(Immunogenic, levels=c("CD8", "negative")), y = StdResid, fill = StdResid)) +
  geom_col() +
  facet_wrap(~reorder(Cluster, -StdResid), nrow = 1) +
  geom_hline(yintercept = c(-2, 2), linetype = "dashed", color = "red") +
  labs(title = "Standardized residuals by supertype and group",
       y = "Standardized residual") +
  scale_fill_gradient2(midpoint = 0) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8,face = "bold"),
    strip.text   = element_text(angle = 90, hjust = 0))
```

### Heatmap
```{r}
# Reorder data_plot based on clusters
ordered_indices <- order(new_data_rearranged$pred)
data_plot_ordered <- new_data_rearranged[ordered_indices,]
col_cluster_annotations <- data_plot_ordered$pred
data_plot_ordered <- data_plot_ordered[,1:7]
data_plot_ordered <- t(data_plot_ordered)

# Prepare annotations based on reordered clusters
col_cluster_annotations <- as.data.frame(col_cluster_annotations)
row.names(col_cluster_annotations) <- colnames(data_plot_ordered)
col_cluster_annotations$col_cluster_annotations <- as.factor(col_cluster_annotations$col_cluster_annotations)

colors <- colorRampPalette(RColorBrewer::brewer.pal(11,"BrBG"))(256)
cluster_colors <- c("1" = "#CC6677", "2" = "#661100", "3" = "#999933", "4" = "#DDCC77", "5" = "#117733", "6" = "#88CCEE", "7" = "#332288", "8" = "black")

pheatmap(data_plot_ordered, cluster_rows=FALSE, cluster_cols=FALSE, show_colnames=FALSE, color = colors, annotation_col = col_cluster_annotations, annotation_colors = list(col_cluster_annotations = cluster_colors))

pdf("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Subcellular_localization/HiTide_data/Pheatmap.pdf")
pheatmap(data_plot_ordered, cluster_rows=FALSE, cluster_cols=FALSE, show_colnames=FALSE, color = colors, annotation_col = col_cluster_annotations, annotation_colors = list(col_cluster_annotations = cluster_colors))
dev.off()
```

## Poisson regression
```{r}
colnames(table) <- c("Cluster", "Immunogenic", "Counts")
fit <- glm(Counts ~ Cluster * Immunogenic,
           data = table,
           family = "poisson")

summary(fit)
```

## Pairwise fisher exact tests

# C1
```{r}
p_values <- matrix(NA, nrow=7, ncol=2)
table_total <- as.data.frame(table(new_data_rearranged$immunogenic))
table_total$Cluster <- "All"
table_total <- table_total[,c(3, 1, 2)]
colnames(table_total) <- c("Cluster", "Immunogenic", "Counts")
print(table_total)

## DO one as a sanity check
table_lim <- table[which(table$Cluster==1),]
  table_diff <- table_total
  table_diff$Counts <- table_diff$Counts - table_lim$Counts
  table_diff <- rbind(table_diff, table_lim)
  df <- table_diff %>%
      # Convert long data into contigency table
      pivot_wider(names_from = Cluster,   
                  values_from = Counts,    
                  values_fill = 0 ) %>%
      column_to_rownames("Immunogenic") %>%
      as.matrix()
  df
  print(table_total) #Validates that I did not change it for next iterations
  fisher.test(df, simulate.p.value = TRUE)

for (i in 1:7){
  table_lim <- table[which(table$Cluster==i),]
  table_diff <- table_total
  table_diff$Counts <- table_diff$Counts - table_lim$Counts
  table_diff <- rbind(table_diff, table_lim)
  df <- table_diff %>%
      # Convert long data into contigency table
      pivot_wider(names_from = Cluster,   
                  values_from = Counts,    
                  values_fill = 0 ) %>%
      column_to_rownames("Immunogenic") %>%
      as.matrix()
  stats <- fisher.test(df, simulate.p.value = TRUE)
  p_values[i, 1] <- i
  p_values[i,2] <- stats$p.value
}

p.adjust(p_values[,2], method = "bonferroni")
```
