---
title: "2024_6_28_Expanded_checkpoint_molecules"
output: html_document
date: '2024-06_28'
---

**Purpose:** Assess association of immune status and metastasis with immune checkpoint and T cell activity expression

**Methods:**

* Plot expression levels of markers of T cell activity and exhaustion

**Results:**

```{r, warning=FALSE, message=FALSE}
library(maftools)
library(stringr)
library(UpSetR)
library(plyr)
library(tidyverse)
library(viridis)
library(patchwork)
library(circlize)
library(networkD3)
library(ggraph)
library(igraph)
library(tidyverse)
library(RColorBrewer)
library(reshape2)
library(data.table)
library(plyr)
library(DGEobj.utils)
library(edgeR)
library(pheatmap)
library(dunn.test)
```


## Add clinical annotations
```{r}
## Read in clinical data
clinical_data <- as.data.frame(read.csv("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Metadata/Metadata_annotations_all_Mayo_samples_immune_binary.csv"))
clinical_data <- clinical_data[which(clinical_data$Dataset=="Old"),]

emily_data <- as.data.frame(read.table("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Metadata/Emily_clinical_data_for_confirmation.txt", header=TRUE))
emily_data <- emily_data[match(clinical_data$Tumor_Sample_Barcode, emily_data$Tumor_Sample_Barcode),]
identical(emily_data$Tumor_Sample_Barcode, clinical_data$Tumor_Sample_Barcode)
identical(emily_data$immune_binary_star, clinical_data$immune_binary)
```

### Read in RNA data
```{r}
expression_TPM <- as.data.frame(read.csv("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/RNA_seq/star/all_samples_tpm.counts.txt", row.names = 1))
```

### Match clinical data

```{r}
expression_TPM <- expression_TPM[,which(colnames(expression_TPM) %in% clinical_data$RNA_ID)]
clinical_data <- clinical_data[match(colnames(expression_TPM), clinical_data$RNA_ID),]
identical(colnames(expression_TPM), clinical_data$RNA_ID)
table(clinical_data$Immunosuppression)

PD_L1 <- expression_TPM[which(row.names(expression_TPM) %in% c("CD274")),]
CTLA4 <- expression_TPM[which(row.names(expression_TPM) %in% c("CTLA4")),]
LAG3 <- expression_TPM[which(row.names(expression_TPM) %in% c("LAG3")),]
PD1 <- expression_TPM[which(row.names(expression_TPM) %in% c("PDCD1")),]
GZMB <- expression_TPM[which(row.names(expression_TPM) %in% c("GZMB")),]
PRF1 <- expression_TPM[which(row.names(expression_TPM) %in% c("PRF1")),]
IL10 <- expression_TPM[which(row.names(expression_TPM) %in% c("IL10")),]
TGFB <- expression_TPM[which(row.names(expression_TPM) %in% c("TGFB1")),]
IFNG <- expression_TPM[which(row.names(expression_TPM) %in% c("IFNG")),]

clinical_data$PDL1 <- t(PD_L1)
clinical_data$CTLA4 <- t(CTLA4)
clinical_data$LAG3 <- t(LAG3)
clinical_data$PD1 <- t(PD1)
clinical_data$GZMB <- t(GZMB)
clinical_data$PRF1 <- t(PRF1)
clinical_data$IL10 <- t(IL10)
clinical_data$TGFB <- t(TGFB)
clinical_data$IFNG <- t(IFNG)
```



### Boxplots with jitter for each comparison
```{r}
## CTLA4
ggplot(clinical_data, aes(x=immune_binary, y= CTLA4)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.1) +theme_minimal()
pdf("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Immune_deconvolution/Functional_markers/CTLA4_boxplot_immune_high_low.pdf")
ggplot(clinical_data, aes(x=immune_binary, y=CTLA4)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.1) + theme_minimal()
dev.off()

kruskal.test(as.numeric(CTLA4) ~ immune_binary, clinical_data)
dunn <- dunn.test(as.numeric(clinical_data$CTLA4), g=clinical_data$immune_binary, method="bh", altp=TRUE)
dunn
```

```{r}
## PDL1
ggplot(clinical_data, aes(x=immune_binary, y= PDL1)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.1) +theme_minimal()
pdf("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Immune_deconvolution/Functional_markers/PDL1_boxplot_immune_high_low..pdf")
ggplot(clinical_data, aes(x=immune_binary, y=PDL1)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.1) + theme_minimal()
dev.off()

kruskal.test(as.numeric(PDL1) ~ immune_binary, clinical_data)
dunn <- dunn.test(as.numeric(clinical_data$PDL1), g=clinical_data$immune_binary, method="bh", altp=TRUE)
dunn
```

```{r}
## LAG3
ggplot(clinical_data, aes(x=immune_binary, y= LAG3)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.1) +theme_minimal()
pdf("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Immune_deconvolution/Functional_markers/LAG3_boxplot_immune_high_low..pdf")
ggplot(clinical_data, aes(x=immune_binary, y=LAG3)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.1) + theme_minimal()
dev.off()

kruskal.test(as.numeric(LAG3) ~ immune_binary, clinical_data)
dunn <- dunn.test(as.numeric(clinical_data$LAG3), g=clinical_data$immune_binary, method="bh", altp=TRUE)
dunn
```

```{r}
## PD1
ggplot(clinical_data, aes(x=immune_binary, y= PD1)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.1) +theme_minimal()
pdf("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Immune_deconvolution/Functional_markers/PD1_boxplot_immune_high_low..pdf")
ggplot(clinical_data, aes(x=immune_binary, y=PD1)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.1) + theme_minimal()
dev.off()

kruskal.test(as.numeric(PD1) ~ immune_binary, clinical_data)
dunn <- dunn.test(as.numeric(clinical_data$PD1), g=clinical_data$immune_binary, method="bh", altp=TRUE)
dunn
```

```{r}
## PRF1
ggplot(clinical_data, aes(x=immune_binary, y= PRF1)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.1) +theme_minimal()
pdf("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Immune_deconvolution/Functional_markers/PRF1_boxplot_immune_high_low..pdf")
ggplot(clinical_data, aes(x=immune_binary, y=PRF1)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.1) + theme_minimal()
dev.off()

kruskal.test(as.numeric(PRF1) ~ immune_binary, clinical_data)
dunn <- dunn.test(as.numeric(clinical_data$PRF1), g=clinical_data$immune_binary, method="bh", altp=TRUE)
dunn
```


```{r}
## GZMB
ggplot(clinical_data, aes(x=immune_binary, y= GZMB)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.1) +theme_minimal()
pdf("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Immune_deconvolution/Functional_markers/GZMB_boxplot_immune_high_low..pdf")
ggplot(clinical_data, aes(x=immune_binary, y=GZMB)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.1) + theme_minimal()
dev.off()

kruskal.test(as.numeric(GZMB) ~ immune_binary, clinical_data)
dunn <- dunn.test(as.numeric(clinical_data$GZMB), g=clinical_data$immune_binary, method="bh", altp=TRUE)
dunn
```

```{r}
ggplot(clinical_data, aes(x=immune_binary, y= IL10)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.1) +theme_minimal()
ggplot(clinical_data, aes(x=immune_binary, y= TGFB+IL10)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.1) +theme_minimal()
ggplot(clinical_data, aes(x=immune_binary, y= IFNG)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.1) +theme_minimal()
```


**Interpretation:**

* There appears to be a consistent trend of highest expression in immunocompetent and metastasized patients 

**Future steps:**

* Consider returning to a more formal differential expression analysis

