---
title: "2023_1_10_Demographic_data_split_immune_met"
output: html_document
date: '2024-01-10'
---

**Original:** "1.0 Demographics/2024_7_2_Demographic_data_summary_exclude_Crohns_patient"

**Purpose:** 

1) To systematically walk through and understand the demographics of the available dataset and how that relates to immunosuppression AND metastasis


**Methods:**

* Evaluate associations of demographics with immunosuppression AND metastasis


**Results:**

```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(maftools)
library(plyr)
library(reshape2)
library(stringr)
library(DGEobj.utils)
library(dunn.test)
```


# Part 1: Investigate demographic data vs. Immunosuppression

### Read in clinical data
```{r}
clin_data <- "~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Metadata/Metadata_annotations_all_Mayo_samples.csv"
clinical_data <- as.data.frame(read.csv(clin_data))
clinical_data <- clinical_data[which(clinical_data$Include_in_study == "Yes"),]
clinical_data <- clinical_data[which(clinical_data$Dataset == "Old"),]
clinical_data <- clinical_data[-which(is.na(clinical_data$Tumor_Sample_Barcode)),]

proportion_binding <- as.data.frame(read.csv("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Metadata/Metadata_annotations_all_Mayo_samples_immune_binary.csv"))
proportion_binding_matched <- proportion_binding[match(clinical_data$Tumor_Sample_Barcode, proportion_binding$Tumor_Sample_Barcode),]
clinical_data$immune_binary <- proportion_binding_matched$immune_binary
clinical_data[which(is.na(clinical_data$immune_binary) & clinical_data$Immunosuppression=="No"),18] <- "No RNA"
clinical_data[which(is.na(clinical_data$immune_binary) & clinical_data$Immunosuppression=="Yes"),18] <- "Suppressed"
clinical_data <- clinical_data[which(clinical_data$Dataset=="Old"),]
table(clinical_data$immune_binary)
```

### Immunosuppressive cause

```{r}
table(clinical_data$Immunosuppression_cause)
```

### Age vs. Immunosuppression:
```{r}
## All
ggplot(clinical_data, aes(x=factor(immune_binary, levels=c("No RNA", "high", "low", "Suppressed")), y=Age)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + 
theme_minimal()
kruskal.test(Age~ immune_binary, clinical_data)
dunn.test(clinical_data$Age, clinical_data$immune_binary, method="bh", altp="TRUE")

pdf("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Demographic Figures/Age_Immune_binary.pdf")
ggplot(clinical_data, aes(x=factor(immune_binary, levels=c("No RNA", "high", "low", "Suppressed")), y=Age)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.05) + 
theme_minimal()
dev.off()
```

### Sex vs. Immunosuppression:
```{r}
#All
table(clinical_data$immune_binary, clinical_data$Gender)

table <- as.data.frame(table(clinical_data$immune_binary, clinical_data$Gender))
table$Freq[c(1,5)] <- table$Freq[c(1,5)]/sum(table$Freq[c(1,5)])
table$Freq[c(2,6)] <- table$Freq[c(2,6)]/sum(table$Freq[c(2,6)])
table$Freq[c(3,7)] <- table$Freq[c(3,7)]/sum(table$Freq[c(3,7)])
table$Freq[c(4,8)] <- table$Freq[c(4,8)]/sum(table$Freq[c(4,8)])
print(table$Freq)

ggplot(table, aes(x= factor(Var1, levels=c("No RNA", "high", "low", "Suppressed")), y = Freq, fill= Var2)) + geom_bar(stat="identity")

table_stats <- table(clinical_data$immune_binary, clinical_data$Gender)
fisher.test(table_stats)

pdf("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Demographic Figures/Sex_Immune_binary.pdf")
ggplot(table, aes(x= factor(Var1, levels=c("No RNA", "high", "low", "Suppressed")), y = Freq, fill= Var2)) + geom_bar(stat="identity") + theme_minimal()
dev.off()
```

### Stage vs. Immunosuppression:
```{r}
#All
table <- as.data.frame(table(clinical_data$immune_binary, clinical_data$Stage))
table$Freq[c(1,5,9)] <- table$Freq[c(1,5,9)]/sum(table$Freq[c(1,5,9)])
table$Freq[c(2,6,10)] <- table$Freq[c(2,6,10)]/sum(table$Freq[c(2,6,10)])
table$Freq[c(3,7,11)] <- table$Freq[c(3,7,11)]/sum(table$Freq[c(3,7,11)])
table$Freq[c(4,8,12)] <- table$Freq[c(4,8,12)]/sum(table$Freq[c(4,8,12)])
print(table$Freq)

ggplot(table, aes(x= factor(Var1, levels=c("No RNA", "high", "low", "Suppressed")), y = Freq, fill= Var2)) + geom_bar(stat="identity")

table_stats <- table(clinical_data$immune_binary, clinical_data$Stage)
fisher.test(table_stats)

pdf("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Demographic Figures/Stage_Immune_binary.pdf")
ggplot(table, aes(x= factor(Var1, levels=c("No RNA", "high", "low", "Suppressed")), y = Freq, fill= Var2)) + geom_bar(stat="identity")+ theme_minimal()
dev.off()
```

### Met vs. Immunosuppression:
```{r}
#All
table <- as.data.frame(table(clinical_data$immune_binary, clinical_data$Met))
table$Freq[c(1,5)] <- table$Freq[c(1,5)]/sum(table$Freq[c(1,5)])
table$Freq[c(2,6)] <- table$Freq[c(2,6)]/sum(table$Freq[c(2,6)])
table$Freq[c(3,7)] <- table$Freq[c(3,7)]/sum(table$Freq[c(3,7)])
table$Freq[c(4,8)] <- table$Freq[c(4,8)]/sum(table$Freq[c(4,8)])
print(table$Freq)

ggplot(table, aes(x= factor(Var1, levels=c("No RNA", "high", "low", "Suppressed")), y = Freq, fill= Var2)) + geom_bar(stat="identity")

table_stats <- table(clinical_data$immune_binary, clinical_data$Met)
fisher.test(table_stats)

pdf("~/3.0 Hasting Research/Immunoediting_human_work_with_new_samples/Demographic Figures/Met_Immune_binary.pdf")
ggplot(table, aes(x= factor(Var1, levels=c("No RNA", "high", "low", "Suppressed")), y = Freq, fill= Var2)) + geom_bar(stat="identity") + theme_minimal()
dev.off()
```





**Interpretation:**

* Difference in age by immunosuppression status

* Slight difference in stage for RNA subset only

* No difference in sex or met status for any group

**Future steps:**

* Finalize figure for manuscript